Estimating Carrying Capacity for Chinook Parr using Quantile Random Forest Models

### 2020-06-11 analysis/paper/paper.Rmd

If a species' carrying capacity is defined or constrained, at least in part, by the habitat in which it lives, then illuminating statistically how such habitat impacts carrying capacity can lead to understanding how a species interacts with its environment. This understanding could be of crucial importance in the realm of conservation, when dealing with an endangered or threatened species, but species/habitat interactions are a core element of ecological studies more generally. We have demonstrated how a quantile regression approach, coupled with a random forest framework, can be used to estimate these relationships, and predict an habitat's capacity. As large ecological datasets become more accessible, and the ability to measure large swaths of habitat more feasible, we believe this approach has many potential applications, from the North American breeding bird survey to groundfish trawl surveys. The framework could also be applied to depleted, non-migratory, and isolated populations (e.g. desert pupfish *Cyprinodon macularius*) to identify limiting factors in populations and/or determine whether a given habitat patch could support a viable population if limiting factors were addressed. Capacity estimates could also be used to potential translocation sites to determine if given sites can support an abundance considered viable before investing in translocation efforts.  ***other potential examples (coral reefs? turtles?)***

>>>>

If a species' carrying capacity is defined or constrained, at least in part, by the habitat in which it lives, then illuminating statistically how such habitat impacts carrying capacity can lead to understanding how a species interacts with its environment. This understanding could be of crucial importance in the realm of conservation, when dealing with an endangered or threatened species, but species/habitat interactions are a core element of ecological studies more generally. We have demonstrated how a quantile regression approach, coupled with a random forest framework, can be used to estimate these relationships, and predict an habitat's capacity. As large ecological datasets become more accessible, and the ability to measure large swaths of habitat more feasible, we believe this approach has many potential applications, from the North American breeding bird survey [@Sauer2017] to groundfish trawl surveys [@Bradburn2011]. The framework could also be applied to depleted, non-migratory, and isolated populations (e.g. desert pupfish *Cyprinodon macularius*) to identify limiting factors in populations and/or determine whether a given habitat patch could support a viable population if limiting factors were addressed. Capacity estimates could also be used to potential translocation sites to determine if given sites can support an abundance considered viable before investing in translocation efforts.  ***other potential examples (coral reefs? turtles?)***
